---
title: "Introduction to distanceto"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Overview

The `distanceto` package is designed to quickly sample distances from points
features to other vector layers. Normally the approach for calculating distance
to (something) involves generating distance surfaces using raster based approaches
eg. `raster::distance` or `gdal_proximity` and subsequently point sampling these
surfaces. Since raster based approaches are a costly method that frequently leads
to memory issues or long and slow run times with high resolution data or large
study sites, we have opted to compute these distances using vector based
approaches. As a helper, there's a decidedly low-res raster based approach for
visually inspecting your region's distance surface. But the workhorse is
`distance_to`.


The `distanceto` package provides two functions:

* `distance_to`
* `distance_raster`


# Example: `distance_to`

```{r setup}
library(distanceto)
library(sf)
```


## Longlat/unprojected
```{r}
# Load nc data
nc <- st_read(system.file("shapes/sids.shp", package="spData"))
st_crs(nc) <- "+proj=longlat +datum=NAD27"

# Set number of sampling points
npts <- 1e3

# Sample points in nc
ncpts <- st_sample(nc, npts)

# Select first 5 of nc
ncsub <- nc[1:5,]

# Measure distance from ncpts to first 5 of nc
dist <- distance_to(ncpts, ncsub, measure = 'geodesic')

# or add to ncpts
ncpts$dist <- dist

head(dist, 30)
hist(dist)
```


## Projected

```{r}
# Load seine data
data('seine', package = 'spData')

# Buffer seine by 1000 metres
bufseine <- st_buffer(seine, 1000)

# Set number of sampling points
npts <- 1e2

# Sample points within buffer seine
seinepts <- st_sample(bufseine, npts)

# Measure distance from seine points to seine
dist <- distance_to(seinepts, seine)

# or add to seine points
seinepts$dist <- dist

head(dist, 30)
hist(dist)
```



# Example: `distance_raster`

```{r}
library(raster)

rdist <- distance_raster(seine, 1e4)

plot(rdist)
```
